Plot curve fits

### 2023-01-04 Working Rmds/plot_fits.Rmd
Split analyses for SD imputation and plotting into separate Rmd documents.

---
title: "<i>invivoPKfit</i> and <i>CvTdb</i> figures"
author: "Christopher Cook, John Wambaugh, Caroline Ring"
date: "11/15/2022"
output:
  html_document: default
  word_document: default
---

@@

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("invivopkfit")
# Clear memory:
rm(list=ls())
```

Read in CvTdb data.

```{r, load_cvt, eval = TRUE}
cvt.data <- read.csv("inst/ext/cvtdb_invivopkfit_04262022-2.csv")
```

A couple of compounds had no CAS in CvTdb. Upon double checking, they should not have been added. So remove them. (CLR 2022-11-15: it doesn't look like there are any of these)

```{r, cvt_remove_nocas}
cvt.data <- cvt.data %>%
  filter(analyte_name_original != "Methemoglobin") %>%
  filter(analyte_name_original != "N-desmethyltamoxifen")
```

Fix a few compound names.
```{r, cvt_fix_cpd_names}
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "Methyleugonal "] <- "Methyleugonal"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "benzo[a]pyrene"] <- "benzo(a)pyrene"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "2,4-D"] <- "2,4-Dichlorophenoxyacetic acid"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "Diclofenac"] <- "Hexobarbital"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "Pentadecafluorooctanoic acid"] <- "Perfluorooctanoic acid"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "3−hydroxybenzo[a]pyrene"] <- "3−hydroxybenzo(a)pyrene"
```

if fk_reference_document is NULL, make fk_reference_document equal fk_extraction document. (CLR 2022-11-15: it doesn't look like there are any of these)

```{r}
cvt.data$fk_reference_document_id[cvt.data$fk_reference_document_id == "NULL"] <- 
  cvt.data$fk_extraction_document_id[cvt.data$fk_reference_document_id == "NULL"]
```

```{r}
### this is basically my debug line,
# where I can focus on one problematic compound 
# and see what's going wrong
# data <- data %>% filter(dsstox_casrn == "68291-97-4" | dsstox_casrn == "82626-48-0")
# data <- data %>% filter(dsstox_casrn == "123-91-1")
```

Calculate a limit of quantification (90% of lowest observation)

```{r, calc_loq, eval = TRUE}
cvt.data <- estimate_loq(cvt.data)
```

Run a one-compartment fit on the data, and write out the results. Timestamp the file name with the current date and time in format YYYY_MM_DD_hh_mm_ss.

Fit the 1-compartment model.

```{r, run_invivoPKfit_1comp}

run_timestamp_1comp <- format(Sys.time(), "%Y_%m_%d_%H_%M_%S")

### do 1-Compartment fit
system.time(
    PK.fit.table.1comp <- fit_all(
      data.set = cvt.data,
      model = "1compartment",
      modelfun = "analytic",
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = TRUE
  )
)

write.csv(PK.fit.table.1comp,
          paste("inst/ext/PK_fit_table_1comp_",
                                    run_timestamp_1comp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)

Fit the 2-compartment model.

```{r, run_invivoPKfit_2comp}

run_timestamp_2comp <- format(Sys.time(), "%Y_%m_%d_%H_%M_%S")

### do 2-Compartment fit
system.time(
    PK.fit.table.2comp <- fit_all(
      data.set = cvt.data,
      model = "2compartment",
      modelfun = "analytic",
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = TRUE
    )
)

write.csv(PK.fit.table.2comp, paste("inst/ext/PK_fit_table_2comp_",
                                    run_timestamp_2comp,
                                    ".csv",
                                    sep=""), row.names = FALSE)

Fit the flat model.

```{r, run_invivoPKfit_flat}
run_timestamp_flat <- format(Sys.time(), "%Y_%m_%d_%H_%M_%S")

system.time(
    PK.fit.table.flat <- fit_all(
      data.set = cvt.data,
      model = "flat",
      modelfun = "analytic",
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = TRUE
    )
)

write.csv(PK.fit.table.flat, paste("inst/ext/PK_fit_table_flat_",
                                    run_timestamp_flat,
                                    ".csv",
                                    sep=""), row.names = FALSE)
```

@@

```{r}
DF <- preprocess_data(data.set = cvt.data,
                       ratio.data.to.dose = 1,
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = FALSE)
```

@@

```{r, merge_pkfits, eval=FALSE}
all.fit.table <- merge_model_fits(list("1compartment"=pk_output_1comp,
                                       "2compartment"=pk_output_2comp,
                                       "flat"=pk_output_flat))
write.csv(all.fit.table, paste("inst/ext/all.fit.table.",THIS.DATE,".csv",sep=""), row.names = FALSE)
```

```{r, make_rdata, eval = FALSE}

### this block takes the csv files in read_data
# and saves them as Rdata files

### save data
paste("inst/ext/PK.fit.table.1comp.",THIS.DATE,".RData",sep="")
save(pk_output_1comp, file = paste("figures_rdata/pk_output_1comp_",THIS.DATE,".RData",sep=""))
save(pk_output_2comp, file = paste("figures_rdata/pk_output_2comp_",THIS.DATE,".RData",sep=""))
save(pk_output_flat, file = paste("figures_rdata/pk_output_flat_",THIS.DATE,".RData",sep=""))
save(raw_1comp, file = paste("figures_rdata/raw_1comp_",THIS.DATE,".RData",sep=""))
save(raw_2comp, file = paste("figures_rdata/raw_2comp_",THIS.DATE,".RData",sep=""))
save(raw_flat, file = paste("figures_rdata/raw_flat_",THIS.DATE,".RData",sep=""))
save(all.fit.table, file = paste("figures_rdata/all_fit_table_",THIS.DATE,".RData",sep=""))
save(pk_lombardo, file = paste("figures_rdata/pk_lombardo",THIS.DATE,".RData",sep=""))
```

```{r, make_invivoPKfit_plots, eval = TRUE}

### this block produces the pdf files of the model fits
# plotted on top of the original data

### 1 compartment

# pk_output_1comp_03012022 <- pk_output_1comp_03012022 %>%
#   filter(Compound == "tamoxifen") %>%
#   filter(Data.Analyzed != "Joint Analysis")

# raw_1comp_03012022 <- raw_1comp_03012022 %>%
#   filter(Compound == "tamoxifen")

junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_1comp),
                      data.set = as.data.table(raw_1comp),
                      model = "1compartment")

### 2 compartment
junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_2comp),
                      data.set = as.data.table(raw_2comp),
                      model = "2compartment")

junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_flat),
                      data.set = as.data.table(raw_flat),
                      model = "flat")

```{r, make_cp_rdata, eval = FALSE}

### this block creates a data.frame of predicted concentrations
# using the estimated parameters and the observed conc-time points

######################

###########
### this is 1-compartment data
###########

###################
###################

### subset to only actual model outputs and discard "Joint Analysis" rows
pk_data1 <- pk_output_1comp %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis") %>% 
  filter(Species != "hamster")

raw_data1 <- dplyr::select(raw_1comp, -c("kelim", 
                                                  "Vdist", 
                                                  "Fgutabs", 
                                                  "kgutabs")) %>% 
  filter(Species != "hamster")

###################
###################

raw_data1$Reference <- as.character(raw_data1$Reference)

full_data_1 <- left_join(raw_data1, pk_data1, by = c("Compound", "CAS", "Reference"))
full_data_1 <- full_data_1 %>% filter(AIC != Inf)

### function that finds cp
find_cp_1 <- function(full_data_1) {plyr::adply(full_data_1, 1, mutate, cp = cp_1comp(time = Time,
                                                                                      params = full_data_1[1, ],
                                                                                      dose = Dose,
                                                                                      iv.dose = iv))}

### have to split data frame into ones with unique parameter values because cp_1comp takes constants
split_df_1 <- split(full_data_1, list(full_data_1$Compound,
                                      full_data_1$Reference,
                                      full_data_1$CAS,
                                      full_data_1$Species.x),
                    drop = TRUE)

cp_df_1 <- lapply(split_df_1, find_cp_1)

cp_df_1 <- do.call(rbind, cp_df_1)

### nomralize cp by Value
cp_df_1$cp_norm <- cp_df_1$cp / cp_df_1$Value
cp_df_1$type <- "1-Compartment"

###########
### this is 2-compartment data
###########

#######################
#######################

### subset to only actual model outputs and discard "Joint Analysis" rows
pk_data2 <- pk_output_2comp %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

raw_data2 <- dplyr::select(raw_2comp, -c("kelim", 
                                                  "Vdist", 
                                                  "Fgutabs", 
                                                  "kgutabs",
                                                  "V1",
                                                  "Ralphatokelim",
                                                  "Fbetaofalpha"))

########################
########################

raw_data2$Reference <- as.character(raw_data2$Reference)

full_data_2 <- left_join(raw_data2, pk_data2, by = c("Compound", "CAS", "Reference"))
full_data_2 <- full_data_2 %>% filter(AIC != Inf)
split_df_2 <- split(full_data_2, list(full_data_2$Compound,
                                      full_data_2$Reference,
                                      full_data_2$CAS,
                                      full_data_2$Species.x),
                    drop = TRUE)

### function that finds cp
find_cp_2 <- function(full_data_2) {plyr::adply(full_data_2, 1, mutate, cp = cp_2comp(time = Time,
                                                                                      params = full_data_2[1, ],
                                                                                      dose = Dose,
                                                                                      iv.dose = iv))}
cp_df_2 <- lapply(split_df_2, find_cp_2)

cp_df_2 <- do.call(rbind, cp_df_2)

cp_df_2$cp_norm <- cp_df_2$cp / cp_df_2$Value
cp_df_2$type <- "2-Compartment"

# cp_df_2 <- cp_df_2 %>%
#   filter(cp != 1e-20)

###########
### this is flat data
###########

pk_data_flat <- pk_output_flat %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis") %>% 
  filter(Species != "hamster")

raw_data_flat <- dplyr::select(raw_flat, -c("A")) %>% 
  filter(Species != "hamster")

raw_data_flat$Reference <- as.character(raw_data_flat$Reference)

full_data_flat <- left_join(raw_data_flat, pk_data_flat, by = c("Compound", "CAS", "Reference"))
full_data_flat <- full_data_flat %>% filter(AIC != Inf)
split_df_flat <- split(full_data_flat, list(full_data_flat$Compound,
                                            full_data_flat$Reference,
                                            full_data_flat$CAS,
                                            full_data_flat$Species.x),
                       drop = TRUE)

### function that finds cp
find_cp_flat <- function(full_data_flat) {plyr::adply(full_data_flat, 1, mutate, cp = cp_flat(time = Time,
                                                                                              params = full_data_flat[1, ],
                                                                                              dose = Dose,
                                                                                              iv.dose = iv))}
cp_df_flat <- lapply(split_df_flat, find_cp_flat)

cp_df_flat <- do.call(rbind, cp_df_flat)

cp_df_flat$cp_norm <- cp_df_flat$cp / cp_df_flat$Value
cp_df_flat$type <- "Flat"

#######################
#######################

### make all data.frames have the same column names so they are rbind-able
cp_df_1 <- cp_df_1[, intersect(colnames(cp_df_2), colnames(cp_df_1))]
cp_df_2 <- cp_df_2[, intersect(colnames(cp_df_1), colnames(cp_df_2))]

cp_df <- rbind(cp_df_1, cp_df_2)

cp_df <- cp_df[, intersect(colnames(cp_df_flat), colnames(cp_df))]
cp_df_flat <- cp_df_flat[, intersect(colnames(cp_df), colnames(cp_df))]

cp_df <- rbind(cp_df, cp_df_flat)

cp_df$cp_pred_error <- (cp_df$cp - cp_df$Value) / cp_df$Value

cp_df <- cp_df %>% group_by(Species.x, 
                           Compound,
                           CAS,
                           type) %>%
 mutate(cp_pred_error_range = max(cp_pred_error, na.rm = TRUE) - min(cp_pred_error, na.rm = TRUE))

### here is where you save the Rdata file
save(cp_df, file = paste("figures_rdata/cp_df_",THIS.DATE,".RData",sep=""))
```

```{r, make_cvt_variability_rdata, eval = FALSE}

### this block creates a data.frame that shows the 
# spread of CvT data via across replicate timepoints 
# for a single experiment
# if the data isn't within 2-fold of itself
# is it reasonable to say that a model should fit within 2-fold?

##################
##################

### all the raw data for each model SHOULD be the same
# so only need to use one
data <- as.data.table(raw_1comp)

##################
##################

### function that finds mean conc of each replicate timepoint
find_mean_conc <- function(data) {

  ### only use data with replicate timepoints
  data <- data[ ave(1:nrow(data), data$Time, FUN = length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {

    time_split <- split(data, data$Time)

    ### discard replicate timepoints where all concentrations are NA
    time_split_list <- purrr::discard(time_split, ~all(is.na(.$Value)))

    if(length(time_split_list) == 0) {
      return(NULL)
    } else {

      data <- do.call(rbind, time_split_list)

      ### where some replicate concentration data are NA, set them equal to the LOQ
      data$Value[is.na(data$Value)] <- data$LOQ[is.na(data$Value)]

      ### take the average Value for each set of timepoints
      data <- data %>%
        group_by(Time) %>%
        do(mutate(., mean_conc = mean(.$Value, na.rm = TRUE)))

      data <- data %>%
        group_by(Time) %>%
        do(mutate(., range_conc = max(.$Value) - min(.$Value)))

      ### create normalized conc values by dividing conc by mean_conc
      data$conc_norm <- data$Value/data$mean_conc

      data <- data[order(data$Time), ]

      return(data)
    }
  }
}

### this for loop applies the function 
# at data subsets of compound, data, dose, route, media
data.list <- list()
for(this.compound in unique(data[, Compound])) {
  this.compound.data <- data[Compound == this.compound]

  ### list of columns by which to split data.set
  col_list <- list(this.compound.data$Reference,
                   this.compound.data$Dose,
                   this.compound.data$Route,
                   this.compound.data$Media)

  ### split data.set into list of data.tables
  split_list <- split(this.compound.data, col_list)

  ### remove data.tables with 0 rows
  split_list <- keep(split_list, ~nrow(.) > 0)

  ### use find_mean_conc fn
  mean_conc <- lapply(split_list, find_mean_conc)

  test <- do.call(rbind, mean_conc)

  data.list[[this.compound]] <- test
}

### combine data.tables back into one large data.set
cvt_variability_df <- do.call(rbind, data.list)

### SAVE the data
save(cvt_variability_df, file = paste("figures_rdata/cvt_variability_df_",THIS.DATE,".RData",sep=""))
```

```{r, load_data}

### this block loads the Rdata files,
# as well as some other files written in below code blocks
# this saves time from having to run those below code blocks every time
# these additional files include 'cp_df' and 'cvt_variability'

### load all of this wonderful data
load(paste("figures_rdata/pk_output_1comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/pk_output_2comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/pk_output_flat_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/raw_1comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/raw_2comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/raw_flat_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/cp_df_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/cvt_variability_df_",THIS.DATE,".RData",sep=""))

load(paste("figures_rdata/cp_df_",THIS.DATE,".RData",sep=""))

load("figures_rdata/pk_lombardo_02072022.Rdata")
```

```{r, make_cp_histogram}

### this block makes a histogram of the data output
# from make_cp_rdata
# there's a separate facet for each model type
# only data that were best fitted by that model
# are shown in that facet

##################
##################

pk_data <- cp_df

# cvt_data <- cvt_variability_df_04282022

##################
##################

row.names(pk_data) <- c()
full_data <- pk_data

# cvt_data <- dplyr::select(cvt_data, -c("kelim",
#                                        "Vdist",
#                                        "Fgutabs",
#                                        "kgutabs"))
# 
# ### split pk_data by model so you can properly join with raw data
# model_split <- split(pk_data, pk_data$model)
# 
# cvt_data$Reference <- as.character(cvt_data$Reference)
# 
# cvt_data$type <- "1-Compartment"
# full_data_1 <- left_join(cvt_data, model_split$`1compartment`)
# 
# cvt_data$type <- "2-Compartment"
# full_data_2 <- left_join(cvt_data, model_split$`2compartment`)
# 
# cvt_data$type <- "Flat"
# full_data_flat <- left_join(cvt_data, model_split$`flat`)
# 
# ### put data back together
# full_data <- rbind(full_data_1, full_data_2, full_data_flat)
# 
# full_data <- full_data %>%
#   filter(cp != 0)

full_data$AIC <- round(full_data$AIC, digits = 4)

### only take necessary columns to keep things clean
full_data <- full_data %>%
  distinct(Compound, Reference, Species.x, model, AIC, cp_norm)

### this pivots the data so we have AICs and cp_norms
# for each model row-wise
full_data <- full_data %>%
  group_by(Species.x, Compound, Reference, model) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = model,
              values_from = c(AIC, cp_norm)) %>%
  dplyr::select(-row)

full_data <- as.data.table(full_data)

### make model column equal a specific model based on AIC values
full_data <- full_data %>% mutate(model = case_when(AIC_1compartment < AIC_2compartment &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    AIC_2compartment < AIC_1compartment &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    AIC_flat < AIC_1compartment &
                                                      AIC_flat < AIC_2compartment ~ "Null",

                                                    !is.na(AIC_1compartment) & is.na(AIC_2compartment) &
                                                      is.na(AIC_flat) ~ "1-Compartment",
                                                    !is.na(AIC_2compartment) & is.na(AIC_1compartment) &
                                                      is.na(AIC_flat) ~ "2-Compartment",
                                                    !is.na(AIC_flat) & is.na(AIC_1compartment) &
                                                      is.na(AIC_2compartment) ~ "Null",

                                                    is.na(AIC_1compartment) &
                                                      AIC_flat < AIC_2compartment ~ "Null",
                                                    is.na(AIC_1compartment) &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    is.na(AIC_2compartment) &
                                                      AIC_flat < AIC_1compartment ~ "Null",
                                                    is.na(AIC_2compartment) &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_1compartment < AIC_2compartment ~"1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_2compartment <AIC_1compartment ~"2-Compartment"))

### make cp_norm equal to the cp_norm that corresponds to the best fitting model
full_data <- full_data %>%
  mutate(cp_norm = case_when(model == "1-Compartment" ~ cp_norm_1compartment,
                             model == "2-Compartment" ~ cp_norm_2compartment,
                             model == "Null" ~ cp_norm_flat))

full_data[, model := paste("Model:", model)]

### this creates a new column that says how many unique sets of Compound, Reference,
# and Species correspond to each model
full_data[, chem_ref_spec_count := uniqueN(.SD), .SDcols = c("Compound", 
                                                             "Reference",
                                                             "Species.x"), 
          by = "model"]

full_data[, chem_ref_spec_count := paste("Chemical-Reference-Species:", chem_ref_spec_count)]

### make plot, facet by model and chem_ref_spec_count
p <- ggplot(data = full_data,
            mapping = aes(x = cp_norm)) +
  geom_histogram(binwidth = 0.325) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  # xlim(0, 1e42) +
  # ylim(0, 400) +
  # scale_y_log10()+
  theme(title = element_text(size = 16),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  # scale_x_log10() +
  scale_x_log10(limits = c(1e-3, 1e3), labels = scientific) +
  scale_y_continuous(labels = scientific) +
  theme_bw() +
  labs(x = "Predicted/Observed",
       y = "Number of Observations") +
  facet_wrap(. ~ model + chem_ref_spec_count)

p
```

```{r, make_cp_error_histogram}

### I think this is little outdated but whatever I'm gonna keep it here
# Think is was just plotting Cpred - Cobs

##################
##################

# data <- cp_df_12212021
# data <- cp_df_03012022
data <- cp_df

##################
##################

data <- data %>%
  filter(cp != 1e-20)

data <- data %>%
  distinct(Species.x, Compound, CAS, cp_pred_error_range)

p <- ggplot(data = data,
            mapping = aes(x = cp_pred_error_range)) +
  geom_histogram() +
  scale_x_log10() +
  facet_wrap(. ~type)

p
```

```{r, make_cp_error_versus_dose}

### I think this is little outdated but whatever I'm gonna keep it here

##################
##################

data <- cp_df

##################
##################

data <- data %>%
  filter(type == "1-Compartment")

data <- data %>%
  group_by(Compound, CAS, Dose) %>%
  mutate(average = mean(cp_pred_error))

p <- ggplot(data = data, mapping = aes(x = Dose, y = average)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()

p

```{r, make_cvt_variability_histogram}

### this block makes the histogram
# of cvt variability using the data.frame
# made from make_cvt_variability_rdata

#################
#################

data <- cvt_variability_df
# data <- cvt_variability_df_12212021

#################
#################

p <- ggplot(data, aes(x = conc_norm)) +
  geom_histogram(binwidth = 0.25) +
  xlim(c(-0.5, 4)) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  theme_bw() +
  labs(x = "Normalized concentration",
       y = "Number of observations")
p
```

```{r, make_density_plot}

### this block makes a density plot with cvt variabity
# and normalized predicted concentrations
# it's basically a combination of the other two histograms
# expect here we limit cp observations those that have replicate timepoint
# data, so it matches the cvt variability data, thus making it 
# comparable

##################
##################

# pk_data <- cp_df_03012022
pk_data <- cp_df

cvt_data <- cvt_variability_df

##################
##################

### i hate row names
row.names(pk_data) <- c()

### making these null so I don't have to deal
# with duplicate .x and .y's
# only need the ones from pk_data
cvt_data <- dplyr::select(cvt_data, -c("kelim", 
                                       "Vdist", 
                                       "Fgutabs", 
                                       "kgutabs"))

### split pk_data by model so you can properly join with raw data
model_split <- split(pk_data, pk_data$model)

### References were of different classes for some reason
# so had to change
cvt_data$Reference <- as.character(cvt_data$Reference)

cvt_data$type <- "1-Compartment"
full_data_1 <- left_join(cvt_data, model_split$`1compartment`)

cvt_data$type <- "2-Compartment"
full_data_2 <- left_join(cvt_data, model_split$`2compartment`)

cvt_data$type <- "Flat"
full_data_flat <- left_join(cvt_data, model_split$`flat`)

### put data back together
full_data <- rbind(full_data_1, full_data_2, full_data_flat)

### not sure why there were 0s,
# but they were screwing things up
full_data <- full_data %>%
  filter(cp != 0)

full_data$AIC <- round(full_data$AIC, digits = 4)

full_data <- full_data %>%
  distinct(Compound, Reference, model, Species.x, conc_norm, AIC, cp_norm)

### 
### 
full_data <- full_data %>%
  group_by(Species.x, Compound, Reference, model) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = model,
              values_from = c(AIC, cp_norm)) %>%
  dplyr::select(-row)

full_data <- full_data %>% mutate(model = case_when(AIC_1compartment < AIC_2compartment & 
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    AIC_2compartment < AIC_1compartment &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    AIC_flat < AIC_1compartment &
                                                      AIC_flat < AIC_2compartment ~ "Null",

                                                    !is.na(AIC_1compartment) & is.na(AIC_2compartment) &
                                                      is.na(AIC_flat) ~ "1-Compartment",
                                                    !is.na(AIC_2compartment) & is.na(AIC_1compartment) & 
                                                      is.na(AIC_flat) ~ "2-Compartment",
                                                    !is.na(AIC_flat) & is.na(AIC_1compartment) & 
                                                      is.na(AIC_2compartment) ~ "Null",

                                                    is.na(AIC_1compartment) &
                                                      AIC_flat < AIC_2compartment ~ "Null",
                                                    is.na(AIC_1compartment) &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    is.na(AIC_2compartment) &
                                                      AIC_flat < AIC_1compartment ~ "Null",
                                                    is.na(AIC_2compartment) &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_1compartment < AIC_2compartment ~"1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_2compartment <AIC_1compartment ~"2-Compartment"))

full_data <- full_data %>%
  mutate(cp_norm = case_when(model == "1-Compartment" ~ cp_norm_1compartment,
                             model == "2-Compartment" ~ cp_norm_2compartment,
                             model == "Null" ~ cp_norm_flat))

### remove super outlier data from triclosan, for visualization purposes
full_data <- full_data %>%
  filter(cp_norm > 1e-10 & cp_norm < 1e8) # also high one for the heck of it

length(unique(full_data$Compound[full_data$model == "1-Compartment"]))
length(unique(full_data$Compound[full_data$model == "2-Compartment"]))
length(unique(full_data$Compound[full_data$model == "Null"]))

null_vec <- as.vector(unique(full_data$Compound[full_data$model == "Null"]))

full_data <- as.data.table(full_data)

full_data[, model := paste("Model:", model)]

full_data[, chem_ref_spec_count := uniqueN(.SD), .SDcols = c("Compound", 
                                                             "Reference",
                                                             "Species.x"), 
          by = "model"]

full_data[, chem_ref_spec_count := paste("Chemical-Reference-Species:", chem_ref_spec_count)]

# full_data <- full_data %>%
#   mutate(cp_norm = case_when(model == "1comparment" ~ cp_norm_1compartment,
#                              model == "2compartment" ~ cp_norm_2compartment))

# full_data$cp_norm[full_data$model == "2compartment"] <- full_data$cp_norm_2compartment
# full_data$cp_norm[full_data$model == "1compartment"] <- full_data$cp_norm_1compartment

# full_data <- full_data %>%
# mutate(across(everything(), ~ifelse(. == "", NA, as.character(.))))

# filter(Compound != "triclosan")
# filter(Compound != "tamoxifen")

###########
###########
###########

split_full_data <- split(full_data, full_data$model)

full_data_1 <- split_full_data$`Model: 1-Compartment`

### Find number of points in each plot panel for 1-compartment
full_data_1_center <- full_data_1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_1_center <- full_data_1_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_1_center <- nrow(full_data_1_center) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_green <- full_data_1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_1_green <- full_data_1_green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_1_green <- nrow(full_data_1_green) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_gray <- full_data_1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_1_gray <- full_data_1_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_1_gray <- nrow(full_data_1_gray) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_tan <- full_data_1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_1_tan <- full_data_1_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_1_tan <- nrow(full_data_1_tan) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

###########
###########

full_data_2 <- split_full_data$`Model: 2-Compartment`

### Find number of points in each plot panel for 2-compartment
full_data_2_center <- full_data_2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_2_center <- full_data_2_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_2_center <- nrow(full_data_2_center) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_green <- full_data_2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_2_green <- full_data_2_green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_2_green <- nrow(full_data_2_green) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_gray <- full_data_2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_2_gray <- full_data_2_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_2_gray <- nrow(full_data_2_gray) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_tan <- full_data_2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_2_tan <- full_data_2_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_2_tan <- nrow(full_data_2_tan) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

###########
###########

full_data_flat <- split_full_data$`Model: Null`

### Find number of points in each plot panel for 2-compartment
full_data_flat_center <- full_data_flat %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_flat_center <- full_data_flat_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_flat_center <- nrow(full_data_flat_center) / nrow(full_data_flat %>% filter(!is.na(cp_norm)))

>>>>

---
title: "Plot curve fits"
author: "Caroline Ring"
date: "2023-01-04"
output: html_document
---

@@

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("../invivopkfit")
```

@@

```{r}
cvt <- preprocess_data(data.set = cvt_data,
                       names_list =list(
                       "Compound_Dosed" = "studies.test_substance_name_original",
                              "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
                              "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
                              "Compound_Analyzed" = "series.analyte_name_original",
                              "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
                              "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
                      "Reference" = "documents_reference.id",
                      "Extraction" = "documents_extraction.id",
                      "Species" = "subjects.species",
                      "Weight" ="subjects.weight_kg",
                      "Weight.Units" = NULL,
                      "Dose" = "correct.dose_level_normalized",
                      "Dose.Units" = NULL,
                      "Time" = "conc_time_values.time_hr",
                      "Time.Units" = NULL,
                      "Media" = "series.conc_medium_normalized",
                      "Value" = "conc_time_values.conc",
                      "Value.Units" = NULL,
                      "Route" = "studies.administration_route_normalized",
                      "LOQ" = "series.loq_normalized",
                      "Subject" = "subjects.id",
                              "N_Subjects" = "series.n_subjects_in_series",
                              "Value_SD" = "conc_time_values.conc_sd"),
                    defaults_list =   list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "h",
                      "Value.Units" = "mg/L"),
      ratio_conc_to_dose = 1,
                    calc_loq_factor = 0.45,
                    routes_keep = c("po", "iv"),
                    media_keep = c("blood", "plasma"),
      suppress.messages = FALSE)
```

@@
